---
title: "MATH 208 - Final Project"
output:
  pdf_document:
    df_print: paged
---

```{r,message=FALSE}
library(tidyverse)
library(ggmosaic)
library(lubridate)
library(gridExtra)
library(knitr)
cpu_gpu_data <- read_csv("chip_dataset.csv")
cpu_gpu_data <- cpu_gpu_data[1:11] %>% as_tibble()
```

# Data preprocessing

```{r}
# find the missing value in each column (except for Release date column)
cpu_gpu_data %>% summarise_all(list(~sum(is.na(.)))) %>% 
 pivot_longer(cols=everything(),names_to = "Variable") %>% 
 filter(value>0)
```

## Remove rows that have missing value

```{r}

# create a function to drop the NA values according to column specified by the passing value

drop_na_row <-  function(data, colname){
  na_rows <- which(is.na(data[, c(colname)]))
  if (length(na_rows) == 0) {
    return(data)
  }
  clean_data <- data[-na_rows,]
  return(clean_data)
}

# create a function to remove all the NA in the tibble
drop_data_na <- function(data){
  tmp <- c()
  summary_data <- is.na(data) %>% apply(1,sum)
  for(rown in seq_along(rownames(data))){
    if(summary_data[[rown]] != 0){
      tmp <- append(tmp,rown)
    }
  }
  data[-tmp,]
}

cpu_gpu_data_clean <- drop_data_na(cpu_gpu_data)

cpu_gpu_data_clean %>% summarise_all(list(~sum(is.na(.)))) %>% 
 pivot_longer(cols=everything(),names_to = "Variable") %>% 
 filter(value>0)

```

------------------------------------------------------------------------

# Task 1

## Part a

### Process Size (nm)

```{r}
# numerical summary for Process Size

cpu_gpu_data_ps_clean <- drop_na_row(cpu_gpu_data, "Process Size (nm)")


cpu_gpu_data_ps_clean %>% group_by(Type) %>% summarise(Avg = mean(`Process Size (nm)`), 
 Med = median(`Process Size (nm)`),
 '25%ile' = quantile(`Process Size (nm)`,0.25),
 '75%ile' = quantile(`Process Size (nm)`,0.75),
 StD = sd(`Process Size (nm)`),
 IQR = IQR(`Process Size (nm)`)
 
 )
```

```{r}
# graphical summary for Process Size

ggplot(cpu_gpu_data_ps_clean, aes(x=`Process Size (nm)`,group=Type,fill=Type)) +
geom_histogram(bins=25,col="black",position = "dodge") +
  labs(title = "Process Size by type")

ggplot(cpu_gpu_data_ps_clean,aes(x=Type,y=`Process Size (nm)`,fill=Type)) +
stat_boxplot(geom="errorbar",width=0.25) + geom_boxplot() +
ylab("Process Size (nm)") +
  labs(title = "Process Size by type")

ggplot(cpu_gpu_data_ps_clean,aes(x=`Process Size (nm)`,fill=Type)) +
geom_density(alpha=0.5)+ xlab("Process Size (nm)") +
  labs(title = "Process Size by type")




```

In terms of process size in nanometer, the performance characteristics of CPU and GPU differ very slightly. While CPUs do have a higher lower bound performance and slightly higher average compared to the CPU.

### TDP (W)

```{r}
# numerical summary for TDP

cpu_gpu_data_TDP_clean <- drop_na_row(cpu_gpu_data, "TDP (W)")

cpu_gpu_data_TDP_clean %>% group_by(Type) %>% summarise(Avg = mean(`TDP (W)`), 
 Med = median(`TDP (W)`),
 '25%ile' = quantile(`TDP (W)`,0.25),
 '75%ile' = quantile(`TDP (W)`,0.75),
 StD = sd(`TDP (W)`),
 IQR = IQR(`TDP (W)`)
 )
```

```{r}
# graphical summary for TDP

ggplot(cpu_gpu_data_TDP_clean, aes(x=`TDP (W)`,group=Type,fill=Type)) +
geom_histogram(bins=25,col="black",position = "dodge") +
  labs(title = "TDP by type")

ggplot(cpu_gpu_data_TDP_clean,aes(x=Type,y=`TDP (W)`,fill=Type)) +
stat_boxplot(geom="errorbar",width=0.25) + geom_boxplot() +
ylab("TDP (W)") +
  labs(title = "TDP by type")

ggplot(cpu_gpu_data_TDP_clean,aes(x=`TDP (W)`,fill=Type)) +
geom_density(alpha=0.5)+ xlab("TDP (W)") +
  labs(title = "TDP by type")




```

There are some distinction between CPU and GPU's Thermal Design Parameter (TDP) in Watt. On average, the performance of CPU is higher than GPU by nearly twelve Watts; it also has a higher upper bound (75% quartile) than the GPU by 21 Watts. However, GPU has a higher lower bound (25% quartile) and median value than CPU.

### Die Size (mm\^2)

```{r}
# numerical summary for Die size

cpu_gpu_Die_data_clean <- drop_na_row(cpu_gpu_data, "Die Size (mm^2)")

cpu_gpu_Die_data_clean %>% group_by(Type) %>% summarise(Avg = mean(`Die Size (mm^2)`), 
 Med = median(`Die Size (mm^2)`),
 '25%ile' = quantile(`Die Size (mm^2)`,0.25),
 '75%ile' = quantile(`Die Size (mm^2)`,0.75),
 StD = sd(`Die Size (mm^2)`),
 IQR = IQR(`Die Size (mm^2)`)
 )
```

```{r}
# graphical summary for Die size
ggplot(cpu_gpu_Die_data_clean, aes(x=`Die Size (mm^2)`,group=Type,fill=Type)) +
  geom_histogram(bins=25,col="black",position = "dodge") +
  labs(title = "Die size by type")

ggplot(cpu_gpu_Die_data_clean,aes(x=Type,y=`Die Size (mm^2)`,fill=Type)) +
  stat_boxplot(geom="errorbar",width=0.25) + geom_boxplot() +
  ylab("Die Size (mm^2)") +
  labs(title = "Die size by type")

ggplot(cpu_gpu_Die_data_clean,aes(x=`Die Size (mm^2)`,fill=Type)) +
  geom_density(alpha=0.5)+ xlab("Die Size (mm^2)") +
  labs(title = "Die size by type")
```

Concerning die size, there are some difference between CPU and GPU performance. The median and 25% quartile are the same. However, the contrast between CPU and GPU's die size are strikingly huge, with GPU die size about 40 square minimeters higher than the CPU. The upper end of the GPU is also significantly higher (alomost 30 mm\^2) compared to the CPU.

### Transistors (million)

```{r}
# numerical summary for Transistors

cpu_gpu_data_Trans_clean <- drop_na_row(cpu_gpu_data, "Transistors (million)")
cpu_gpu_data_Trans_clean <- subset(cpu_gpu_data_Trans_clean, `Transistors (million)` < 20000)

cpu_gpu_data_Trans_clean %>% group_by(Type) %>% summarise(Avg = mean(`Transistors (million)`), 
                                                    Med = median(`Transistors (million)`),
                                                    '25%ile' = quantile(`Transistors (million)`,0.25),
                                                    '75%ile' = quantile(`Transistors (million)`,0.75),
                                                    StD = sd(`Transistors (million)`),
                                                    IQR = IQR(`Transistors (million)`)
)
```

```{r,warning=FALSE}
# graphical summary for Transistors
ggplot(cpu_gpu_data_Trans_clean, aes(x=`Transistors (million)`,group=Type,fill=Type)) +
  geom_histogram(bins=25,col="black",position = "dodge") +
  labs(title = "Transistors by type")

ggplot(cpu_gpu_data_Trans_clean,aes(x=Type,y=`Transistors (million)`,fill=Type)) +
  stat_boxplot(geom="errorbar",width=0.25) + geom_boxplot(horizontal=TRUE,axes=FALSE,outline=FALSE) +
  ylab("Transistors (million)") +
  labs(title = "Transistors by type") 

ggplot(cpu_gpu_data_Trans_clean,aes(x=`Transistors (million)`,fill=Type)) +
  geom_density(alpha=0.5)+ xlab("Transistors (million)") +
  labs(title = "Transistors by type") + scale_x_log10()
```

In comparison between the number of transistors CPU and GPU, all the statistical figures suggest that the performance of the GPU is much worse than CPU as it uses a lot more transistors.

### Freq (MHz)

```{r}
# numerical summary for Frequency

cpu_gpu_data_Freq_clean <- drop_na_row(cpu_gpu_data, "Freq (MHz)")

cpu_gpu_data_Freq_clean %>% group_by(Type) %>% summarise(Avg = mean(`Freq (MHz)`), 
                                                     Med = median(`Freq (MHz)`),
                                                     '25%ile' = quantile(`Freq (MHz)`,0.25),
                                                     '75%ile' = quantile(`Freq (MHz)`,0.75),
                                                     StD = sd(`Freq (MHz)`),
                                                     IQR = IQR(`Freq (MHz)`)
)
```

```{r}
# graphical summary for Freq
ggplot(cpu_gpu_data_clean, aes(x=`Freq (MHz)`,group=Type,fill=Type)) +
  geom_histogram(bins=25,col="black",position = "dodge") +
  labs(title = "Freq by type")

ggplot(cpu_gpu_data_clean,aes(x=Type,y=`Freq (MHz)`,fill=Type)) +
  stat_boxplot(geom="errorbar",width=0.25) + geom_boxplot() +
  ylab("Freq (MHz)") +
  labs(title = "Freq by type")

ggplot(cpu_gpu_data_clean,aes(x=`Freq (MHz)`,fill=Type)) +
  geom_density(alpha=0.5)+ xlab("Freq (MHz)") +
  labs(title = "Freq by type")
```

When it comes to running frequency in hertz, all the CPU's data are overwhelmingly superior than the GPU performance.

------------------------------------------------------------------------

## b

```{r}
# find the occurrence of different companies in Foundry and Vendor
  cpu_gpu_Foundary_data <- cpu_gpu_data %>% group_by(Foundry) %>% count(name = "counts")

  cpu_gpu_Vendor_data <- cpu_gpu_data %>% group_by(Vendor) %>% count(name = "counts")
```

### collapse foundries with smaller counts into a single level

```{r}
cpu_gpu_data_collapse <- cpu_gpu_data
for(i in seq_along(cpu_gpu_data_collapse$Foundry)){
  if(cpu_gpu_data_collapse$Foundry[i] == "NEC" |
     cpu_gpu_data_collapse$Foundry[i] == "Renesas" |
     cpu_gpu_data_collapse$Foundry[i] == "Samsung" |
     cpu_gpu_data_collapse$Foundry[i] == "Sony" |
     cpu_gpu_data_collapse$Foundry[i] == "UMC"){
    cpu_gpu_data_collapse$Foundry[i] = "other"
  }
}

cpu_gpu_Foundary_data_collapse <- cpu_gpu_data_collapse %>% group_by(Foundry) %>% count(name = "counts")
```

### All vendors

```{r,message=FALSE}
# numerical summary of all vendors
cpu_gpu_data_collapse %>% group_by(Vendor,Type,Foundry) %>% 
  summarise(count = n()) %>% 
  mutate(prop_by_vendor_type = count/sum(count)) 
```

```{r,message=FALSE}
# graphical summary of all vendors
cpu_gpu_data_collapse %>% group_by(Type) %>%
  ggplot(aes(x=Foundry,fill=Foundry)) + 
  geom_bar() + scale_fill_viridis_d() + labs(title="Foundry preference counts of All vendors") + facet_wrap(~Type)


cpu_gpu_data_collapse %>% group_by(Type,Foundry) %>% 
  summarise(count = n()) %>% 
  mutate(prop_by_Type = count/sum(count)) %>%
  ggplot(aes(x="",y=prop_by_Type,
             fill=Foundry)) + 
  geom_bar(stat="identity") + coord_polar("y",start=0) +
  theme(axis.title.y =element_blank(),
        axis.text.y =element_blank(),
        axis.ticks.y=element_blank()) +
  scale_fill_viridis_d() +
  ylab("Proportion") +
  labs(title="Foundry preference proportion of All vendors") + 
  facet_wrap(~Type)

cpu_gpu_data_collapse  %>%
  ggplot() + geom_mosaic(aes(x=product(Foundry,Vendor),
 fill=Vendor))+ 
 theme(axis.title.x =element_blank(),axis.text.x =element_blank(),
 axis.ticks.x=element_blank()) +
  labs(title = " Association between the number of processors released by the vendors and the
foundries by type") + 
  facet_wrap(~Type)
```

More than half of of customers chose the Intel CPU, followed by unknown brand. For GPU, more than 79% of the buyers preferred the TSMC foundry, while other brands are equally preferred by the customers.

### AMD

```{r,message=FALSE}
# numerical summary of AMD
cpu_gpu_data_collapse %>% filter(Vendor == "AMD") %>% group_by(Type,Foundry) %>% 
  summarise(count = n()) %>% 
  mutate(prop_by_Type = count/sum(count)) 
```

```{r,message=FALSE}
# graphical summary of AMD
cpu_gpu_data_collapse %>% filter(Vendor == "AMD") %>% group_by(Type) %>%
ggplot(aes(x=Foundry,fill=Foundry)) + 
 geom_bar() + scale_fill_viridis_d() + labs(title="Foundry preference counts of AMD") + facet_wrap(~Type)


cpu_gpu_data_collapse %>% filter(Vendor == "AMD") %>% group_by(Type,Foundry) %>% 
  summarise(count = n()) %>% 
  mutate(prop_by_Type = count/sum(count)) %>%
  ggplot(aes(x="",y=prop_by_Type,
 fill=Foundry)) + 
 geom_bar(stat="identity") + coord_polar("y",start=0) +
 theme(axis.title.y =element_blank(),
 axis.text.y =element_blank(),
 axis.ticks.y=element_blank()) +
 scale_fill_viridis_d() +
  ylab("Proportion") +
  labs(title="Foundry preference proportion of AMD") + 
  facet_wrap(~Type)
```

For machineries of the AMD, most customers (75%) preferred TSMC foundary while the most popular CPU of AMD are produced by unknown foundary.

### ATI

```{r,message=FALSE}
# numerical summary of ATI
cpu_gpu_data_collapse %>% filter(Vendor == "ATI") %>% group_by(Type,Foundry) %>% 
  summarise(count = n()) %>% 
  mutate(prop_by_Type = count/sum(count)) 
```

```{r,message=FALSE}
# graphical summary of ATI
cpu_gpu_data_collapse %>% filter(Vendor == "ATI") %>% group_by(Type) %>%
  ggplot(aes(x=Foundry,fill=Foundry)) + 
  geom_bar() + scale_fill_viridis_d() + labs(title="Foundry preference counts of ATI") + facet_wrap(~Type)


cpu_gpu_data_collapse %>% filter(Vendor == "ATI") %>% group_by(Type,Foundry) %>% 
  summarise(count = n()) %>% 
  mutate(prop_by_Type = count/sum(count)) %>%
  ggplot(aes(x="",y=prop_by_Type,
             fill=Foundry)) + 
  geom_bar(stat="identity") + coord_polar("y",start=0) +
  theme(axis.title.y =element_blank(),
        axis.text.y =element_blank(),
        axis.ticks.y=element_blank()) +
  scale_fill_viridis_d() +
  ylab("Proportion") +
  labs(title="Foundry preference proportion of ATI") + 
  facet_wrap(~Type)
```

### Intel

```{r,message=FALSE}
# numerical summary of Intel
cpu_gpu_data_collapse %>% filter(Vendor == "Intel") %>% group_by(Type,Foundry) %>% 
  summarise(count = n()) %>% 
  mutate(prop_by_Type = count/sum(count)) 
```

```{r,message=FALSE}
# graphical summary of Intel
cpu_gpu_data_collapse %>% filter(Vendor == "Intel") %>% group_by(Type) %>%
  ggplot(aes(x=Foundry,fill=Foundry)) + 
  geom_bar() + scale_fill_viridis_d() + labs(title="Foundry preference counts of Intel") + facet_wrap(~Type)


cpu_gpu_data_collapse %>% filter(Vendor == "Intel") %>% group_by(Type,Foundry) %>% 
  summarise(count = n()) %>% 
  mutate(prop_by_Type = count/sum(count)) %>%
  ggplot(aes(x="",y=prop_by_Type,
             fill=Foundry)) + 
  geom_bar(stat="identity") + coord_polar("y",start=0) +
  theme(axis.title.y =element_blank(),
        axis.text.y =element_blank(),
        axis.ticks.y=element_blank()) +
  scale_fill_viridis_d() +
  ylab("Proportion") +
  labs(title="Foundry preference proportion of Intel") + 
  facet_wrap(~Type)
```

### NVIDIA

```{r,message=FALSE}
# numerical summary of NVIDIA
cpu_gpu_data_collapse %>% filter(Vendor == "NVIDIA") %>% group_by(Type,Foundry) %>% 
  summarise(count = n()) %>% 
  mutate(prop_by_Type = count/sum(count)) 
```

```{r,message=FALSE}
# graphical summary of NVIDIA
cpu_gpu_data_collapse %>% filter(Vendor == "NVIDIA") %>% group_by(Type) %>%
  ggplot(aes(x=Foundry,fill=Foundry)) + 
  geom_bar() + scale_fill_viridis_d() + labs(title="Foundry preference counts of NVIDIA") + facet_wrap(~Type)


cpu_gpu_data_collapse %>% filter(Vendor == "NVIDIA") %>% group_by(Type,Foundry) %>% 
  summarise(count = n()) %>% 
  mutate(prop_by_Type = count/sum(count)) %>%
  ggplot(aes(x="",y=prop_by_Type,
             fill=Foundry)) + 
  geom_bar(stat="identity") + coord_polar("y",start=0) +
  theme(axis.title.y =element_blank(),
        axis.text.y =element_blank(),
        axis.ticks.y=element_blank()) +
  scale_fill_viridis_d() +
  ylab("Proportion") +
  labs(title="Foundry preference proportion of NVIDIA") + 
  facet_wrap(~Type)
```

For NVIDIA, TSMC foundry accounts for about 90% of the production, while other foundries account for the same.

### other vendors

```{r,message=FALSE}
# numerical summary of other vendors
cpu_gpu_data_collapse %>% filter(Vendor == "Other") %>% group_by(Type,Foundry) %>% 
  summarise(count = n()) %>% 
  mutate(prop_by_Type = count/sum(count)) 
```

```{r,message=FALSE}
# graphical summary of other vendors
cpu_gpu_data_collapse %>% filter(Vendor == "Other") %>% group_by(Type) %>%
  ggplot(aes(x=Foundry,fill=Foundry)) + 
  geom_bar() + scale_fill_viridis_d() + labs(title="Foundry preference counts of Other vendors") + facet_wrap(~Type)


cpu_gpu_data_collapse %>% filter(Vendor == "Other") %>% group_by(Type,Foundry) %>% 
  summarise(count = n()) %>% 
  mutate(prop_by_Type = count/sum(count)) %>%
  ggplot(aes(x="",y=prop_by_Type,
             fill=Foundry)) + 
  geom_bar(stat="identity") + coord_polar("y",start=0) +
  theme(axis.title.y =element_blank(),
        axis.text.y =element_blank(),
        axis.ticks.y=element_blank()) +
  scale_fill_viridis_d() +
  ylab("Proportion") +
  labs(title="Foundry preference proportion of Other vendors") + 
  facet_wrap(~Type)
```

------------------------------------------------------------------------

## c

```{r}
# numerically shows the association between Die Size and TDP
cpu_gpu_data_clean %>% group_by(Type) %>% 
  summarise(Cor_TDP_DieSize = cor(`TDP (W)`,`Die Size (mm^2)`))
```

```{r,message=FALSE}
# graphically shows the association between Die Size and TDP
ggplot(cpu_gpu_data_clean,aes(x=`TDP (W)`,y=`Die Size (mm^2)`,col=Type)) +
  geom_point() + 
 geom_smooth(method="lm",col="black") +
  labs(title = " Association between Die Size and TDP by Type") + 
  facet_wrap(~Type)
```

The association of between die size and TDP is quite different between CPU and GPU in general, with GPU has a much higher correlation compared to CPU.

------------------------------------------------------------------------

# Task 2

## a

### Remove rows have missing release date

```{r}
missing_date <- function(value){
  ifelse(value == "NaT", TRUE, FALSE)
}

# number of missing release date rows
cpu_gpu_data_collapse$`Release Date` %>% missing_date() %>% sum()
```

```{r}
drop_missing_date <- function(data){
  tmp <- c()
  summary_data <- missing_date(data$`Release Date`) %>% as.numeric()
  for(rown in seq_along(rownames(data))){
    if(summary_data[[rown]] != 0){
      tmp <- append(tmp,rown)
    }
  }
  data[-tmp,]
}



cpu_gpu_data_clean_date <- drop_missing_date(cpu_gpu_data_clean)

# number of missing release date rows
cpu_gpu_data_clean_date$`Release Date` %>% missing_date() %>% sum()
```

```{r}
# change the data type of  column "Release Date" from char to Date
 cpu_gpu_data_clean_date <- cpu_gpu_data_clean_date %>%
  mutate(`Release Date` = (ymd(cpu_gpu_data_clean_date$`Release Date`)))
```

### Foundry

```{r,message=FALSE}
# numerical summary
cpu_gpu_data_clean_date %>% mutate(year_only = year(`Release Date`)) %>% 
 group_by(year_only,Foundry) %>%
 summarise(count=n())%>%kable(.)
```

```{r,message=FALSE,warning=FALSE}
# graphical summary
cpu_gpu_data_clean_date %>% mutate(year_only = year(`Release Date`)) %>% 
 group_by(year_only,Foundry) %>%
 summarise(count=n()) %>%
  ggplot(aes(x=year_only, y= count)) + 
  geom_line() + 
 geom_point(col="red") + 
 labs(x="Year",y="Number of processors relased",
      title = "Number of processors relased by year of Foundry" ) + 
  geom_smooth() +
 theme(axis.text.x = element_text(angle = 45)) + 
  facet_wrap(~Foundry)

# Observation of trend of the number of processors released by year of foundry 

# GF: there is a strong positive association between year and number of processors from 2014 to 2017 and a strong negative association from 2017 t0 2021

# Intel: starting from 2000, there is a weak positive association between year and number of processors released peaking at 100 in 2010. Then, there is a negative association 
# between the processors and year from 2010 to 2021

# Other: the number of processors released stays constant throughout 10 years at 0

# TSMC: the trend of processors released by year is similar to Intel. The positive association between number of processors and year is stronger than Intel in first 10 years.
# In 2013, the processors released is around 170 which is the peak of five different categories of company, then it decreases as shown in the graph until 2020

# Unknown: the trend of unknown is that the data fluctuates over 20 years which generally increase in the first 10 years and decrease in the latter 10 years
```

### Vendor

```{r,message=FALSE}
# numerical summary
cpu_gpu_data_clean_date %>% mutate(year_only = year(`Release Date`)) %>% 
 group_by(year_only,Vendor) %>%
 summarise(count=n())
```

```{r,message=FALSE,warning=FALSE}
# graphical summary
cpu_gpu_data_clean_date %>% mutate(year_only = year(`Release Date`)) %>% 
 group_by(year_only,Vendor) %>%
 summarise(count=n()) %>%
  ggplot(aes(x=year_only, y= count)) + 
  geom_line() + 
 geom_point(col="red") + 
 labs(x="Year",y="Number of processors relased",
      title = "Number of processors relased by year of Vendor" ) + 
  geom_smooth() +
 theme(axis.text.x = element_text(angle = 45)) + 
  facet_wrap(~Vendor)

# Observation of the trend of processors released by year of Vendor

# AMD: there is huge variation between years and there is a increase from the first 10 years, and decrease in the last 10 years. But the  association is not strong as 
# the data varies heavily in the last 10 years

# ATI: the graph shows the data of processors built by ATI in the first 13 years. there is a strong positive association from 2000 to 2007 and the figure decrease from 50 to 0 in 
# the last 6 years

# Intel: there is a obvious trend of number of processors built by Intel. The data increases from 2000 to 2010(strong positive association) and decrease from 2020 to 2020( 
# strong negative association)

# NVIDIA: generally the trend of NVIDIA is that the data increase slightly from 2000 to 2010, but at 2005 and 2007, there are two outliers of the overall trend

# Other: similarly to the trend in foundry, the data of other stays constant at around 10 from 2000 to 2003. No other trend is found
```

------------------------------------------------------------------------

## b

```{r,message=FALSE}
# numerical summary
cpu_gpu_data_clean_date <- cpu_gpu_data_clean_date %>% mutate(year_only = year(`Release Date`))
  
cpu_gpu_data_clean_date %>% 
  group_by(year_only,Type) %>% 
  summarise(mean_transistor_by_year_type = mean(`Transistors (million)`)) %>%
  filter(year_only %% 2 == 0)
```

```{r}
# simulate Moore's law
tmp1 <- seq(1:11)
tmp2 <- seq(1:11)
for(i in seq(1:11)){
  tmp1[i] <- 38*2^(i-1)
  tmp2[i] <- 27.21429*2^(i-1)
}
Moore <- tibble(year_only = seq(2000,2020,2), CPU = tmp1, GPU = tmp2) %>%
  pivot_longer(cols=c(CPU,GPU),names_to = "Type",values_to = "mean_transistor_by_year_type")

Moore
```

```{r,message=FALSE,warning=FALSE}
# graphical summary
p1 = cpu_gpu_data_clean_date %>% 
  group_by(year_only,Type) %>% 
  summarise(mean_transistor_by_year_type = mean(`Transistors (million)`)) %>%
  ggplot(aes(x=year_only, y= mean_transistor_by_year_type)) + 
  geom_line() + 
  geom_point(col="red") + 
  labs(x="Year",y="Number of transistor relased",
       title = "Number of transistor relased by every two year" ) + 
  geom_smooth() +
  theme(axis.text.x = element_text(angle = 45)) +
  scale_x_continuous(breaks = seq(2000,2020,2)) + 
  facet_wrap(~Type)

p2 = ggplot(Moore,aes(x=year_only, y= mean_transistor_by_year_type)) + 
  geom_line() + 
  geom_point(col="red") + 
  labs(x="Year",y="Number of transistor relased",
       title = "Number of transistor relased by every two year based on Moore's law" ) + 
  geom_smooth() +
  theme(axis.text.x = element_text(angle = 45)) +
  scale_x_continuous(breaks = seq(2000,2020,2)) + 
  facet_wrap(~Type)

grid.arrange(p1,p2)

# Based on the Moore's Law: he number of transistors per microchip would double every two years. This is generally true for GPU and false for CPU. By observing the numerical value # # # which is computed to following graph. It is obviously to find that number of transistor released every two years increase rapidly which fits the trend given by Moore's law well
# through 20 years. However, the growth of CPU is smaller and the evidence to contradict Moore's law is that from 2018 to 2020, it only increases from 5000 to 10000 but the 
# Moore's law predicts that it will increase from 20000 to 40000.

```
